DeepSphere: a graph-based spherical CNN

### 2019-11-15 main.tex
figure 2: rename
polynomial order P instead of K to avoid confusion with k neighbors

\paragraph{Convolution.} \label{sec:method:convolution}
On Euclidean domains, convolutions are efficiently implemented by sliding a window in the signal domain.
On the sphere however, there is no straightforward way to implement a convolution in the signal domain due to non-uniform samplings.
Convolutions are most often performed in the spectral domain through a spherical harmonic transform (SHT).
That is the approach taken by \citet{cohen2018sphericalcnn} and \citet{esteves2018sphericalcnn}, which has a computational cost of $\bO(n^{3/2})$ on isolatitude samplings (such as the HEALPix and equiangular samplings) and $\bO(n^2)$ in general.
On the other hand, following \citet{defferrard2016graphnn}, graph convolutions can be defined as
\begin{equation} \label{eqn:graph_conv}
	h(\b{L}) \b{f} = \left(\sum_{i=0}^K \alpha_i \b{L}^i\right) \b{f},
\end{equation}
where $K$ is the polynomial order (which corresponds to the filter's size) and $\alpha_i$ are the coefficients to be optimized during training.\footnote{In practice, training with Chebyshev polynomials (instead of monomials) is slightly more stable. We believe it to be due to their orthogonality and uniformity.}
Those convolutions are used by \citet{khasanova2017sphericalcnn} and \citet{perraudin2019deepspherecosmo} and cost $\bO(n)$ operations through a recursive application of $\b{L}$.\footnote{As long as the graph is sparsified such that the number of edges, i.e., the number of non-zeros in $\b{A}$, is proportional to the number of vertices $n$. This can always be done as most weights are very small.}

@@

\paragraph{Theoretical convergence.}
\begin{wrapfigure}{r}{5cm}
	\centering
	\includegraphics[width=\linewidth]{d_iA_i}
	\caption{The patch around the $i$th sampling point.}
\end{wrapfigure}
We extend the work of \citep{belkin2005towards} to a sufficiently regular, deterministic sampling.
Following their setting, we will work with the extended graph Laplacian operator as the linear operator $L_n^t: L^{2}(\mathcal{M}) \rightarrow L^{2}(\mathcal{M})$ such that
\begin{equation} \label{eq:Heat Kernel Graph Laplacian operator}
	 L_n^t f(y) := \frac{1}{n}\sum_{i=1}^{n} e^{ -\frac{\|x_i-y\|^2}{4t}} \left(f(y)-f(x_i)\right).
\end{equation}
This operator extends the graph Laplacian operator previously defined with the weighting scheme \eqnref{weights:belkin} to each point of the the sphere.
% \mdeff{Can we avoid to introduce general manifolds and stay on the sphere?} \nati{yes}
As the radius of the kernel $t$ will be adapted to the number of samples, we also rescale the graph Laplacian operator as $\hat{L}_n^t := \frac{|\mathbb S^2|}{4\pi t^2}L_n^t$.
Hence, we have $\b{L}_n^t \b{f} = T_\V L_n^t f$.
% $\b{\hat L} = \frac{|\mathbb S^2|}{4\pi t^2}\b L_n^t$. \nati{why define this, we are not using it}
Given a sampling $\V$, we define $\sigma_i$ to be the patch of the surface of the sphere corresponding to $x_i$, $A_i$ its corresponding area, and $d_i$ the largest distance between the center $x_i$ and any point on the surface $\sigma_i$.
Define $d^{(n)} := \max_{i=1, \dots, n} d_i$ and $A^{(n)} := \max_{i=1, \dots, n} A_i$.
%Given any rotation $g\in SO(3)$, we define the corresponding rotation operator $\mathcal R_g$ such that for any function $f$ on the sphere
%\mdeff{Already defined in problem formulation.}
%\begin{equation} \label{eq:rotation operator}
%	\mathcal R_g f(x) := f(g^{-1}(x)).
%\end{equation}
\begin{theorem}
	For a sampling $\V$ of the sphere that is equi-area and such that $d^{(n)} \leq \frac{C}{n^\alpha}, \ \alpha\in (0,\linefrac{1}{2}]$, for all $f: \S^2 \rightarrow \R$ Lipschitz with respect to the Euclidean distance in $\R^3$, for all $y\in\S^2$, there exists a sequence $t_n = n^\beta$, $\beta\in\mathbb R$ such that
% 	the rescaled Heat Kernel Graph Laplacian $\frac{|\S^2|}{4\pi t_n^2}\b {L}^t_n$ converges pointwise to the Laplace Beltrami operator on the sphere $\Delta{\S^2}$
	\begin{equation*}
		\lim_{n\to\infty}\hat{L}_n^{t_n}f(y) = \triangle_{\S^2}f(y).
	\end{equation*}
	\label{theo:pointwise convergence for a regular sampling}
\end{theorem}
As a direct implication of this theorem, the rescaled graph Laplacian (the discrete counterpart of $\hat L^t_n$) converges pointwise towards the Laplace-Beltrami operator $\Delta_{\S^2}$ in the sense explained in \citep{belkin2007convergence}.
This is a major step towards equivariance as the Laplace-Beltrami operator does commute with rotation. Based on this property, we show that the graph Laplacian is equivariant to rotation in the limit of infinite sampling.

@@

The equiangular graph is built with \eqnref{weights:khasanova} and $k = 4$ neighbors \citep[following][]{khasanova2017sphericalcnn}.
The HEALPix graph is built with \eqnref{weights:belkin}, $k = 8$, and a kernel width $t$ set to the average of the distances \citep[following][]{perraudin2019deepspherecosmo}.
The NN is made of $5$ graph convolutional layers, each followed by a max pooling layer which down-samples by $4$.
A GAP and a fully connected layer with softmax follow.
The polynomials are all of order $K=3$ and the number of channels per layer is $16, 32, 64, 128, 256$, respectively.
Following \citet{esteves2018sphericalcnn}, the cross-entropy plus a triplet loss is optimized with Adam for 30 epochs on the dataset augmented by 3 random translations.
The learning rate is $5 \cdot 10^{-2}$ and the batch size is 32.

@@

Graphs are built with \eqnref{weights:belkin}, $k = 8, 20, 40$ neighbors, and the corresponding optimal kernel widths $t$ given in \secref{optimal}.
Following \citet{perraudin2019deepspherecosmo}, the NN is made of $5$ graph convolutional layers, each followed by a max pooling layer which down-samples by $4$.
A GAP and a fully connected layer with softmax follow.
The polynomials are all of order $K=4$ and the number of channels per layer is $16, 32, 64, 64, 64$, respectively.
The cross-entropy loss is optimized with Adam for 80 epochs.
The learning rate is $2 \cdot 10^{-4} \cdot 0.999^{\textrm{step}}$ and the batch size is 8.

@@

The graph is built with \eqnref{weights:belkin}, $k = 6$ neighbors, and a kernel width $t$ set to the average of the distances.
Following \citet{jiang2019sphericalcnn}, the NN is an encoder-decoder with skip connections.
Details in \secref{climate:appendix}.
%The encoder is made of $5$ graph convolutional layers, each followed by a \todo{pick-one?} pooling layer which down-samples by $4$.
%the number of channels per layer is $16, 32, 64, 64, 64$, respectively.
%The decoder is made of
The polynomials are all of order $K=3$.
The cross-entropy loss (weighted or non-weighted) is optimized with Adam for 30 epochs.
The learning rate is $1 \cdot 10{-3}$ and the batch size is 64.

@@

The graph is built with \eqnref{weights:belkin}, $k = 5$ neighbors, and a kernel width $t$ set to the average of the distances.
The equivariance property of the resulting graph has not been tested, and we don't expect it to be good due to the heavily non-uniform sampling.
The NN is made of $3$ graph convolutional layers.
The polynomials are all of order $K=0$ or $4$ and the number of channels per layer is $50, 100, 100$, respectively.
For the global regression, a GAP and a fully connected layer follow.
For the dense regression, a graph convolutional layer follows instead.
The MSE loss is optimized with RMSprop for 250 epochs.
The learning rate is $1 \cdot 10^{-3}$ and the batch size is 64.

\begin{table}
    \centering
    \begin{tabular}{
			c
			S[table-format=2.2]
			S[table-format=1.2]
			S[table-format=1.3]
			S[table-format=1.2]
			S[table-format=1.2]
			S[table-format=1.3]
			S[table-format=1.2]
			S[table-format=1.2]
			S[table-format=-1.3]
		}
		\toprule
		& \multicolumn{3}{c}{temp. (from past temp.)} & \multicolumn{3}{c}{day (from temperature)} & \multicolumn{3}{c}{day (from precipitations)} \\
		\cmidrule(lr){2-4} \cmidrule(lr){5-7} \cmidrule(lr){8-10}
		order $K$ & {MSE} & {MAE} & {R2} & {MSE} & {MAE} & {R2} & {MSE} & {MAE} & {R2} \\
		\midrule
		$0$ & 10.88 & 2.42 & 0.896 & 0.10 & 0.10 & 0.882 & 0.58 & 0.42 & -0.980 \\
		%%$1$ &  8.91 & 2.20 & 0.906 & - & - & - & - & - & - \\
		$4$ &  8.20 & 2.11 & 0.919 & 0.05 & 0.05 & 0.969 & 0.50 & 0.18 &  0.597 \\
		%$9$ &  8.38 & 2.12 & 0.915 & - & - & - & - & - & - \\
		\bottomrule
    \end{tabular}
    \caption{
		Prediction results on data from weather stations.
		Structure always improves performance.
	}
    \label{tab:ghcn}
\end{table}

Results are shown in \tabref{ghcn}.
While using a polynomial order $K=0$ is like modeling each time series independently with an MLP, orders $K>0$ integrate neighborhood information.
Results show that using the structure induced by the spherical geometry always yields better performance.

>>>>

\paragraph{Convolution.} \label{sec:method:convolution}
On Euclidean domains, convolutions are efficiently implemented by sliding a window in the signal domain.
On the sphere however, there is no straightforward way to implement a convolution in the signal domain due to non-uniform samplings.
Convolutions are most often performed in the spectral domain through a spherical harmonic transform (SHT).
That is the approach taken by \citet{cohen2018sphericalcnn} and \citet{esteves2018sphericalcnn}, which has a computational cost of $\bO(n^{3/2})$ on isolatitude samplings (such as the HEALPix and equiangular samplings) and $\bO(n^2)$ in general.
On the other hand, following \citet{defferrard2016graphnn}, graph convolutions can be defined as
\begin{equation} \label{eqn:graph_conv}
	h(\b{L}) \b{f} = \left(\sum_{i=0}^P \alpha_i \b{L}^i\right) \b{f},
\end{equation}
where $P$ is the polynomial order (which corresponds to the filter's size) and $\alpha_i$ are the coefficients to be optimized during training.\footnote{In practice, training with Chebyshev polynomials (instead of monomials) is slightly more stable. We believe it to be due to their orthogonality and uniformity.}
Those convolutions are used by \citet{khasanova2017sphericalcnn} and \citet{perraudin2019deepspherecosmo} and cost $\bO(n)$ operations through a recursive application of $\b{L}$.\footnote{As long as the graph is sparsified such that the number of edges, i.e., the number of non-zeros in $\b{A}$, is proportional to the number of vertices $n$. This can always be done as most weights are very small.}

@@

\paragraph{Theoretical convergence.}
\begin{wrapfigure}{r}{5cm}
	\centering
	\includegraphics[width=\linewidth]{patch}
	\caption{The patch around the $i$th sampling point.}
\end{wrapfigure}
We extend the work of \citep{belkin2005towards} to a sufficiently regular, deterministic sampling.
Following their setting, we will work with the extended graph Laplacian operator as the linear operator $L_n^t: L^{2}(\mathcal{M}) \rightarrow L^{2}(\mathcal{M})$ such that
\begin{equation} \label{eq:Heat Kernel Graph Laplacian operator}
	 L_n^t f(y) := \frac{1}{n}\sum_{i=1}^{n} e^{ -\frac{\|x_i-y\|^2}{4t}} \left(f(y)-f(x_i)\right).
\end{equation}
This operator extends the graph Laplacian operator previously defined with the weighting scheme \eqnref{weights:belkin} to each point of the the sphere.
% \mdeff{Can we avoid to introduce general manifolds and stay on the sphere?} \nati{yes}
As the radius of the kernel $t$ will be adapted to the number of samples, we also rescale the graph Laplacian operator as $\hat{L}_n^t := \frac{|\mathbb S^2|}{4\pi t^2}L_n^t$.
Hence, we have $\b{L}_n^t \b{f} = T_\V L_n^t f$.
% $\b{\hat L} = \frac{|\mathbb S^2|}{4\pi t^2}\b L_n^t$. \nati{why define this, we are not using it}
Given a sampling $\V$, we define $\sigma_i$ to be the patch of the surface of the sphere corresponding to $x_i$, $A_i$ its corresponding area, and $d_i$ the largest distance between the center $x_i$ and any point on the surface $\sigma_i$.
Define $d^{(n)} := \max_{i=1, \dots, n} d_i$ and $A^{(n)} := \max_{i=1, \dots, n} A_i$.
%Given any rotation $g\in SO(3)$, we define the corresponding rotation operator $\mathcal R_g$ such that for any function $f$ on the sphere
%\mdeff{Already defined in problem formulation.}
%\begin{equation} \label{eq:rotation operator}
%	\mathcal R_g f(x) := f(g^{-1}(x)).
%\end{equation}
\begin{theorem}
	For a sampling $\V$ of the sphere that is equi-area and such that $d^{(n)} \leq \frac{C}{n^\alpha}, \ \alpha\in (0,\linefrac{1}{2}]$, for all $f: \S^2 \rightarrow \R$ Lipschitz with respect to the Euclidean distance in $\R^3$, for all $y\in\S^2$, there exists a sequence $t_n = n^\beta$, $\beta\in\mathbb R$ such that
% 	the rescaled Heat Kernel Graph Laplacian $\frac{|\S^2|}{4\pi t_n^2}\b {L}^t_n$ converges pointwise to the Laplace Beltrami operator on the sphere $\Delta{\S^2}$
	\begin{equation*}
		\lim_{n\to\infty}\hat{L}_n^{t_n}f(y) = \triangle_{\S^2}f(y).
	\end{equation*}
	\label{theo:pointwise convergence for a regular sampling}
\end{theorem}
As a direct implication of this theorem, the rescaled graph Laplacian (the discrete counterpart of $\hat L^t_n$) converges pointwise towards the Laplace-Beltrami operator $\Delta_{\S^2}$ in the sense explained in \citep{belkin2007convergence}.
This is a major step towards equivariance as the Laplace-Beltrami operator does commute with rotation. Based on this property, we show that the graph Laplacian is equivariant to rotation in the limit of infinite sampling.

@@

The equiangular graph is built with \eqnref{weights:khasanova} and $k = 4$ neighbors \citep[following][]{khasanova2017sphericalcnn}.
The HEALPix graph is built with \eqnref{weights:belkin}, $k = 8$, and a kernel width $t$ set to the average of the distances \citep[following][]{perraudin2019deepspherecosmo}.
The NN is made of $5$ graph convolutional layers, each followed by a max pooling layer which down-samples by $4$.
A GAP and a fully connected layer with softmax follow.
The polynomials are all of order $P=3$ and the number of channels per layer is $16, 32, 64, 128, 256$, respectively.
Following \citet{esteves2018sphericalcnn}, the cross-entropy plus a triplet loss is optimized with Adam for 30 epochs on the dataset augmented by 3 random translations.
The learning rate is $5 \cdot 10^{-2}$ and the batch size is 32.

@@

Graphs are built with \eqnref{weights:belkin}, $k = 8, 20, 40$ neighbors, and the corresponding optimal kernel widths $t$ given in \secref{optimal}.
Following \citet{perraudin2019deepspherecosmo}, the NN is made of $5$ graph convolutional layers, each followed by a max pooling layer which down-samples by $4$.
A GAP and a fully connected layer with softmax follow.
The polynomials are all of order $P=4$ and the number of channels per layer is $16, 32, 64, 64, 64$, respectively.
The cross-entropy loss is optimized with Adam for 80 epochs.
The learning rate is $2 \cdot 10^{-4} \cdot 0.999^{\textrm{step}}$ and the batch size is 8.

@@

The graph is built with \eqnref{weights:belkin}, $k = 6$ neighbors, and a kernel width $t$ set to the average of the distances.
Following \citet{jiang2019sphericalcnn}, the NN is an encoder-decoder with skip connections.
Details in \secref{climate:appendix}.
%The encoder is made of $5$ graph convolutional layers, each followed by a \todo{pick-one?} pooling layer which down-samples by $4$.
%the number of channels per layer is $16, 32, 64, 64, 64$, respectively.
%The decoder is made of
The polynomials are all of order $P=3$.
The cross-entropy loss (weighted or non-weighted) is optimized with Adam for 30 epochs.
The learning rate is $1 \cdot 10{-3}$ and the batch size is 64.

@@

The graph is built with \eqnref{weights:belkin}, $k = 5$ neighbors, and a kernel width $t$ set to the average of the distances.
The equivariance property of the resulting graph has not been tested, and we don't expect it to be good due to the heavily non-uniform sampling.
The NN is made of $3$ graph convolutional layers.
The polynomials are all of order $P=0$ or $4$ and the number of channels per layer is $50, 100, 100$, respectively.
For the global regression, a GAP and a fully connected layer follow.
For the dense regression, a graph convolutional layer follows instead.
The MSE loss is optimized with RMSprop for 250 epochs.
The learning rate is $1 \cdot 10^{-3}$ and the batch size is 64.

\begin{table}
    \centering
    \begin{tabular}{
			c
			S[table-format=2.2]
			S[table-format=1.2]
			S[table-format=1.3]
			S[table-format=1.2]
			S[table-format=1.2]
			S[table-format=1.3]
			S[table-format=1.2]
			S[table-format=1.2]
			S[table-format=-1.3]
		}
		\toprule
		& \multicolumn{3}{c}{temp. (from past temp.)} & \multicolumn{3}{c}{day (from temperature)} & \multicolumn{3}{c}{day (from precipitations)} \\
		\cmidrule(lr){2-4} \cmidrule(lr){5-7} \cmidrule(lr){8-10}
		order $P$ & {MSE} & {MAE} & {R2} & {MSE} & {MAE} & {R2} & {MSE} & {MAE} & {R2} \\
		\midrule
		$0$ & 10.88 & 2.42 & 0.896 & 0.10 & 0.10 & 0.882 & 0.58 & 0.42 & -0.980 \\
		%%$1$ &  8.91 & 2.20 & 0.906 & - & - & - & - & - & - \\
		$4$ &  8.20 & 2.11 & 0.919 & 0.05 & 0.05 & 0.969 & 0.50 & 0.18 &  0.597 \\
		%$9$ &  8.38 & 2.12 & 0.915 & - & - & - & - & - & - \\
		\bottomrule
    \end{tabular}
    \caption{
		Prediction results on data from weather stations.
		Structure always improves performance.
	}
    \label{tab:ghcn}
\end{table}

Results are shown in \tabref{ghcn}.
While using a polynomial order $P=0$ is like modeling each time series independently with an MLP, orders $P>0$ integrate neighborhood information.
Results show that using the structure induced by the spherical geometry always yields better performance.
